---
title: What parameters give a better mileage for a car? Does type of Transmission matter?
author: "Kuber K Gudipati"
date: "Sunday, July 26, 2015"
output: pdf_document
---
###Context
You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:
."Is an automatic or manual transmission better for MPG"
."Quantify the MPG difference between automatic and manual transmissions"

###Data 
```{r results='hide'}
library(ggplot2)
data(mtcars)

mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)

levels(mtcars$am) <- c("Automatic", "Manual"); 
```
###Initial Analysis
Let's go with problem statement and draw a boxplot between mpg and am to see what type of transmission gives a better mileage. See the figures in Appendix.  

It's evdent from the boxplot that Manual Transmission gives better mileage compared to Automatic. Lets draw the conclusion from regression models. Lets go with a liner regression models
### Regression Analysis
```{r, results='hide'}
mpgamModel <-lm(mpg~am, data = mtcars)
summary(mpgamModel)
```
R-squared:  0.3598 means, it only explaind 36% of variance in mpg. There may be  other potential variables affecting mpg. Let's plot with all the variables
```{r, results='hide'}
fullModel <-lm(mpg~., data = mtcars)
summary(fullModel)
```
Based on the above summary (Estimate and t-values), CylinderCount, horsepower, weight and carb2 appear to be significant. Lets do stepwise model and figure out the most significant factors
```{r, results='hide'}
stepModel <-step(fullModel,direction="backward", k=log(nrow(mtcars)))
summary(stepModel)
```

Based  on the above stepwise model, wt + qsec + am seems to be significant on mpg. Let analyze this model

```{r, results='hide'}
model2<-lm(mpg ~ wt + qsec + am, data = mtcars)
summary(model2)
```
Lets compare mpgamModel and model2 and decide which one is better
```{r, results='hide'}
anova(mpgamModel,stepModel)
```
Since the p-vale <.05, model2 seems to give better results. 

###Conclusion
```{r, results='hide'}
summary(model2)
```
It appears that, keeping weight and accelertion (1/4th mile distance) constant, Manual Transmission may give 2.936 more miles per gallon

###Appendix
Boxplot between Manual and Auto Transmission vs mpg

```{r, echo=FALSE}
ggplot(data=mtcars, aes(x=am, y=mpg, fill=am)) + geom_boxplot()
```

Model1: mpg~am
```{r}
mpgamModel <-lm(mpg~am, data = mtcars)
summary(mpgamModel)
```

stepModel:
```{r, results='hide'}
stepModel <-step(fullModel,direction="backward", k=log(nrow(mtcars)))
summary(stepModel)
```
summary of model2:
```{r}
model2<-lm(mpg ~ wt + qsec + am, data = mtcars)
summary(model2)
```
Lets compare mpgamModel and model2 and decide which one is better:
```{r}
anova(mpgamModel,stepModel)
```
Conclusion
```{r}
summary(model2)
```
```{r}
par(mfrow = c(2,2))
plot(model2)
```